Inferring true changes in microbial abundance from taxonomically-biased microbiome measurements

### 2022-01-19 introduction.Rmd
Draft abstract

>>>>

::: {#abstract .rmdnote}

**Abstract:**
To assess how the species in microbial communities vary across conditions, researchers often measure the relative or absolute abundances of all species simultaneously using marker-gene and metagenomic sequencing.
These measurements are taxonomically biased, with some species measured many times more efficiently than others.
Measurement error caused by taxonomic bias is generally ignored in such _differential abundance (DA) analyses_, likely due to a combination of 1) a widespread belief that protocol standardization is sufficient to yield accurate differences between samples and 2) a lack of practical solutions for addressing it.
Using theoretical arguments and analysis of real and simulated experiments, we show that taxonomic bias can create error in DA results that is sufficient to lead to scientifically erroneous conclusions.
This error, however, varies by experimental context and analysis method.
We present six approaches to addressing the error, each suited to different experimental questions and systems.
These approaches provide practical solutions to mitigating the effect of taxonomic bias for a large fraction of DA questions.
If adopted, they may lead to notable improvements in the reproducibility and biological interpretability of the statistical findings of microbiome studies.

:::
